\documentclass[11pt]{article}

\usepackage[utf8]{inputenc}
\usepackage[sfdefault]{roboto}
\usepackage[T1]{fontenc}
\usepackage[left=0.75in,top=0.4in,right=0.75in,bottom=0.6in]{geometry}
\usepackage{array}
\usepackage{tabu}

\usepackage[svgnames]{xcolor}
\definecolor{lightgrey}{HTML}{95A5A6}
\definecolor{grey}{rgb}{0.25,0.25,0.25}
\newcommand{\blackb}{\color{Black} \usefont{OT1}{lmss}{m}{n}}
\newcommand{\lightgreyb}{\color{lightgrey} \usefont{OT1}{lmss}{m}{n}}

\let\bold\textbf
\let\ital\textit
\let\smca\textsc
\let\cour\texttt

\renewcommand{\rmdefault}{Roboto}

%\setfont{subsection}{\usefont{T1}{fvm}{m}{n}}
%\setfont{section}{\usefont{T1}{fvs}{b}{n}\Large}

%\headerfont{Roboto}
%\headerfont[
%  UprightFont=*-Regular,
%  ItalicFont=*-Italic,
%  BoldFont=*-Bold,
%  BoldItalicFont=*-BoldItalic,
%]{Roboto}


\begin{document}
\pagenumbering{gobble}
\begin{center}
  {\fontsize{33}{60}\selectfont {\fontseries{t}\selectfont Ayush} \bold{Bansal}} \\
  {\scriptsize {\color{lightgrey}\smca{Second Year Undergraduate $\cdot$ Computer Science and Engineering}}} \\
  {\footnotesize \ital{\color{lightgrey}Indian Institute of Technology}} \\
  {\scriptsize (+91) 94625-45882 $\mid$ ayushb268@gmail.com}
\end{center}
\rule{\textwidth}{2pt}\vspace{0.2cm}
I am enthusiastic about \bold{Systems} and learning new stuff about \ital{Linux architecture, Programming languages, networking} and \ital{security}. I have experience with cloud computing, ethical hacking and participate in \bold{CTFs} in my free time. I want to pursue a career in research and do a significant contribution in the field of \bold{Computer Science}. \\
\rule{\textwidth}{2pt}
\section*{{\fontseries{t}\selectfont Edu}cation\color{gray}\hrulefill}

\begin{tabu} to \textwidth {| c | l | X[c] | r |}
  \hline
  \bold{Year} & \bold{Degree} & \bold{Institution(Board)} & \bold{CGPA/\%} \\
  \hline
  July'16 -- Present & B.Tech, CSE & Indian Institute of Technology, Kanpur & 8.5/10.0 \\
  2016 & CBSE -- XII & Saint Josephs School, Kota (CBSE) & 95\% \\
  2014 & CBSE -- X & Saint Pauls School, Kota (CBSE) & 94\%\\
  \hline
\end{tabu}

\section*{{\fontseries{t}\selectfont Hon}ors \& Awards\color{gray}\hrulefill}

\begin{tabu} to \textwidth {c X[l] r}
  2016 & \bold{All India Rank 187}, Joint Entrance Exam Advanced, 200,000 candidates & {\footnotesize\fontseries{t}\selectfont\ital{India}} \\
  2016 & \bold{All India Rank 571}, Joint Entrance Exam Mains, 1.5 million candidates & {\footnotesize\fontseries{t}\selectfont\ital{India}} \\
  2016 & \bold{All India Rank 52}, KVPY Scholarship, Indian Institute of Science & {\footnotesize\fontseries{t}\selectfont\ital{Bangalore,India}} \\
  2015 & \bold{Merit Certificate, National Top 1\%}, Indian National Chemistry Olympiad & {\footnotesize\fontseries{t}\selectfont\ital{India}} \\
  2015 & \bold{Merit Certificate, National Top 1\%}, Indian National Astronomy Olympiad & {\footnotesize\fontseries{t}\selectfont\ital{India}}
\end{tabu}

\section*{{\fontseries{t}\selectfont Wor}k Experience\color{gray}\hrulefill}

\begin{tabu} to \textwidth {l X[r]}
  \bold{New York Office, IIT Kanpur} & {\scriptsize\fontseries{t}\selectfont\ital{IIT Kanpur}} \\
  {\footnotesize\color{lightgrey}\smca{Infrastructure Management, Supervisor Prof. Manindra Agarwal}} & {\scriptsize\fontseries{t}\selectfont\ital{May 2017 - Present}} 
\end{tabu}
{\footnotesize\fontseries{t}\selectfont\begin{itemize} % Description(s) of tasks/responsibilities
        \item {Worked on a scalable microservice based web application with an extensive technology stack.}
        \item {Deployed microservices utilizing Docker container, Coreos Container Linux, JenkinsCI, and have maintained the cluster on Kubernetes.}
        \item {Configured and deployed Persistent Volume Storage System.}
        \item {Technologies used: Kubernetes, Scala with Akka, Docker, Nginx, Coreos Applications, Rook with Cephfs, Heketi with Glusterfs.}
        \item {Summer Internship continued as a volunteer and peer mentor.}
\end{itemize}}

\section*{{\fontseries{t}\selectfont Pro}jects\color{gray}\hrulefill}

\begin{tabu} to \textwidth {l X[r]}
  \bold{Implementation of Dynamic Multi-Level Paging} & {\scriptsize\fontseries{t}\selectfont\ital{IIT Kanpur}} \\
  {\footnotesize\color{lightgrey}\smca{Advisor Prof. Debadutta Mishra}} & {\scriptsize\fontseries{t}\selectfont\ital{Oct. 2017 - Present}} 
\end{tabu}
{\footnotesize\fontseries{t}\selectfont\begin{itemize} % Description(s) of tasks/responsibilities
        \item {Improve performance by creating a better \textbf{Virtual Address} organization.}
        \item {Find wastage (internal fragmentation) in page-table pages by studying the page table usage of different applications.}
        \item {Results gathered from application's page-table usage are used to create a better page-table organization.}

\end{itemize}}
\newpage
\noindent\begin{tabu} to \textwidth {l X[r]}
  \bold{Ethical Hacking: Learnt Basic Exploits for participating in CTFs} & {\scriptsize\fontseries{t}\selectfont\ital{IIT Kanpur}} \\
  {\footnotesize\color{lightgrey}\smca{Programming Club}} & {\scriptsize\fontseries{t}\selectfont\ital{Summer'2017}} 
\end{tabu}
{\footnotesize\fontseries{t}\selectfont\begin{itemize} % Description(s) of tasks/responsibilities
        \item {Learnt exploitation techniques like buffer-overflow, integer-overflow, format strings and heap exploits.}
        \item {Used debuggers and knowledge of x86 Assembly language to exploit binary files.}
        \item {Performed various web attacks like sql-injection, XSS and CSRF.}
        \item {Learnt basics of Cryptography.}
\end{itemize}}

\noindent\newline\begin{tabu} to \textwidth {l X[r]}
  \bold{Linux From Scratch: Compiled Linux System from Source and Built a Package Manager} & {\scriptsize\fontseries{t}\selectfont\ital{IIT Kanpur}} \\
  {\footnotesize\color{lightgrey}\smca{Association of Computing Activities}} & {\scriptsize\fontseries{t}\selectfont\ital{Feb. - Apr. 2017}} 
\end{tabu}
{\footnotesize\fontseries{t}\selectfont\begin{itemize} % Description(s) of tasks/responsibilities
        \item {Created a \textbf{Customized Linux} based operating system by compiling all packages from scratch using the host machine's kernel.}
        \item {Built a \textbf{package manager} in python which fetched the dependencies of packages and created a dependency list based on packages installed.}
\end{itemize}}

\section*{{\fontseries{t}\selectfont Rel}evant Coursework\color{gray}\hrulefill}

\hspace{-2.55mm}{\fontsize{11pt}{1.5em}
  \begin{tabular*}{\textwidth}{l l l}
    Introduction to Programming  & Discrete Mathematics & Logic in Computer Science \\
    Abstract Algebra (A$*$) & Data Structures and Algorithms ($i$) & Computer Organization ($i$) \\
    Programming Tools and Techniques ($i$) \\
\end{tabular*}

\vspace{2mm}
\hspace{-7.4mm}\begin{tabular}{l l}
  {\footnotesize\color{lightgrey}A$*$: Grade for Exceptional Performance} & {\footnotesize\color{lightgrey}$i$: In progress}
\end{tabular}
}

\section*{{\fontseries{t}\selectfont Ski}lls\color{gray}\hrulefill}

\begin{tabular} {r l}
  \bold{Languages} & C/C++, Bash, Python, Scala, Golang, Javascript \\
  \bold{Utilities} & Linux shell utilities, Git, Docker, Kubernetes, Akka with Scala, GDB, \LaTeX, Vim \\
  \bold{Operating Systems} & Arch Linux, Ubuntu \\
\end{tabular}

\section*{{\fontseries{t}\selectfont Ext}racurricular Activity\color{gray}\hrulefill}

\begin{tabu} to \textwidth {l X[r]}
  \bold{Leader} {\footnotesize\smca{Google's Developer Students Club, IIT Kanpur}} & {\scriptsize\fontseries{t}\selectfont\ital{July 2017 - Present}} \\
  \bold{Secretary} {\footnotesize\smca{Programming Club, IIT Kanpur}} & {\scriptsize\fontseries{t}\selectfont\ital{Apr. 2017 - Present}} \\
\end{tabu}

\end{document}
